Timely Object Recognition

### 2012-05-31 writeups/nips12/technical.tex
paper

\section{Multi-class Recognition Policy} \label{sec:tech}
\begin{figure}[h!]
\center{\includegraphics[width=0.66\linewidth]
    {../figures/pomdp.png}}
  \caption{Summary of our approach to the problem. Our system has two major parts: (1) selecting an action by predicting its value; (2) updating the belief state with observations resulting from the action.}
  \label{fig:pomdp}
\end{figure}

Our goal is a multi-class recognition policy $\pi$ that takes an image $\mathcal{I}$ and outputs a list of multi-class detection results by running detector and other \emph{actions} sequentially.

@@

We refer to the basis for the decisions made by the decision process as the \emph{state} $b$.
The state includes the currently believed distribution over class presence variables $P(\mathbf{C}) = P(C_1, \dots, C_K)$, where we write $P(C_k)$ to mean $P(C_k=1)$

@@

A recognition \emph{episode} takes an image $\mathcal{I}$ and proceeds from the initial state $b^0$ and action $a^0$ to the next pair $(b^1,a^1)$, and so on until $t$ exceeds $T_d$.
At that point, the policy is terminated, and a new episode can begin on a new image.

\begin{table}[h!]
\centering
\caption{Summary of the notation.}
\label{tab:notation}
\begin{tabular}{|l|l|}
  \hline
  $\mathcal{I}$ & image \\
  $C_k$         & presence of class $k \in \{1,\dots,K\}$ \\ 
  $t$           & time into episode \\ 
  $T_s$, $T_d$  & start and deadline times \\ 
  $b^j$         & belief state at step $j$ \\ 
  $\pi$         & policy function, $b \mapsto a \in \mathcal{A}$ \\
  $\mathcal{A}$ & set of actions $a$\\ 
  \comment{$\mathcal{F}$ & set of featurization actions \\}
  \comment{$\mathcal{L}$ & set of classification actions\\}
  $o_i$         & a real-valued observation upon executing $a_i \in \mathcal{A}$\\
  $\mathcal{O}$ & set of executed actions\\
  $\mathbf{o}$  & set of observations $\{o_i | a_i \in \mathcal{O}\}$\\
  $c(a_i)$        & cost of executing $a_i$, in units of $t$\\
  \hline
\end{tabular}\end{table}

\subsection{Selecting actions based on expected reward} \label{sec:value}
As our goal is to pick actions dynamically, we want a function $V(b,a): B \times \mathcal{A} \mapsto \mathbb{R}$, where $B$ is the space of all possible states, to assign a value to a potential action, given the current state of the decision process.
We can then define the desired policy $\pi$ as simply taking the (untaken) action with the maximum value:
\begin{align}
\pi(b) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} V(b,a_i)
\end{align}

This formulation can represent a closed-loop policy if the observations $\mathbf{o}$ (part of the state $b$) play a role in determining the value of a potential action.

Before proceeding, we note that that although the action space $\mathcal{A}$ is quite manageable, consisting of the detectors and classifier we would like to run on the image, the space of possible states $B$ is infinite.
Therefore we cannot learn a tabular representation of $V(b,a)$, and so use function approximation to represent it \cite{Sutton1998}.
We featurize the state-action pair and assume linear structure:
\begin{align}
V^\pi(b,a_i) = \theta_\pi^\top  \phi(b,a_i)
\end{align}

The policy's performance at time $t$ is determined by the detections that are part of the set of observations $\mathbf{o}$ at the last state $b^j$ before $t$.
Detection results of unobserved classes are an empty set.
Therefore, the final AP vs. Time evaluation of the episode is a function of the history of execution $h^0=b^0,b^1,\dots,b^J$, where $J$ is the last step of the process with $t \le T_d$.
We call the final evaluation function $eval(h^0,T_s,T_d)$, and it is precisely the area under the AP vs. Time curve (normalized by total possible area) between $T_s$ and $T_d$, as determined by the detections in $\mathbf{o^j}$ for all steps $j$ in the episode $h$.

@@

Specifically, as shown in Figure~\ref{fig:rewards}, we define the \emph{reward} of an action as
\begin{align}\label{eq:advanced}
R(b^j,a_i) = \frac{\Delta \text{ap}_i (t_T^j-\frac{1}{2}\Delta t_i)}{(1-\text{ap}^j)t_T^j}
\end{align}
where $t_T^j$ and $\text{ap}^j$ are the time left until deadline and the AP at state $b^j$, and $\Delta t_i$ and $\Delta \text{ap}_i$ are the time taken and AP change produced by the action $a_i$.

@@

We can then represent the final evaluation in terms of individual rewards:
\begin{align}
eval(h^0) = \sum_{j=0}^J R(b_j,a^{j})
\end{align}

And we can represent the expected value of the final evaluation recursively in terms of the value function:
\begin{align} \label{eq:recursive_value}
V(b^j,a_i) = \mathbb{E}_{b^{j+1}} [R(b^j,a_i) + \gamma V(b^{j+1},\pi(b^{j+1}))]
\end{align}
where $\gamma \in [0,1]$ is a \emph{discount} value that can mitigate the effects of increasing state-transition uncertainty over long episodes.

\subsection{Feature representation and learning the policy}
\todo{write out current approach}

Let us take take the feature representation for an action $a_i$ to be $[P(C_k), \, 1]$, where $C_k$ corresponds to the action.
This corresponds to our heuristic value function, with a bias variable.
The feature representation $\phi(b,a_i)$ then is a vector of size $F|\mathcal{A}|$, with $F=2$ for this feature, where all values are $0$ except those corresponding to $a_i$.

The feature representation of the belief state for a given action $a_i$ that corresponds to class $k$ is defined to be
\begin{align}
\phi_k(b) = [P(C_k), \, H(C_k), \, 1]
\end{align}

Our procedure for learning the weights is standard generalized policy iteration \cite{Sutton1998}.
We first initialize the weights $\theta$ to the heuristic value function of simply picking the maximum $P(C_k)$: $\theta_i = [1, \,0, \,0]$.

With this policy, we run $N$ recognition episodes.
From the state-action samples gathered in running the episodes, we formulate a matrix $\Phi$ from the featurizations $\phi(b^j,a_i)$, and a vector $y$ consisting of the individual rewards $R(b^j,a_i)$.

The rewards are computed as the sum of discounted rewards to the end of the episode:
\begin{align}
R(b^j,a_i) = \sum_{i=0}^{J-j} \lambda^i R(b_{j+i},a^{j+i})
\end{align}
Note that with $\lambda=0$, the reward is determined entirely by the actual action taken; with $\lambda=1$, the reward is the sum of all rewards until the end of the episode.

We then solve the system $\Phi \theta = y$ for $\theta$ with Lasso regression.
With the updated weights $\theta$, we run $N$ more episodes and repeat the procedure until convergence.

The parameters of this training procedure are the $\alpha$ weight on the regularization term in the regression and the $\lambda$ discount weight.
We cross-validate for both values.

\subsection{Updating probabilities}
\begin{figure}[h!]
\centering
\includegraphics[width=0.56\linewidth]{../figures/mrf_w_gist.pdf}
\caption{
The MRF inference model used in our system. The shaded nodes represent two actions that have already been taken; their observations have been recorded.
}
\label{fig:model}
\end{figure}

>>>>

\section{Multi-class Recognition Policy} \label{sec:tech}
Our goal is a multi-class recognition policy $\pi$ that takes an image $\mathcal{I}$ and outputs a list of multi-class detection results by running detector and other \emph{actions} sequentially.

@@

We refer to the basis for the decisions made by the decision process as the \emph{state} $s$.
The state includes the currently believed distribution over class presence variables $P(\mathbf{C}) = P(C_1, \dots, C_K)$, where we write $P(C_k)$ to mean $P(C_k=1)$

@@

A recognition \emph{episode} takes an image $\mathcal{I}$ and proceeds from the initial state $s^0$ and action $a^0$ to the next pair $(s^1,a^1)$, and so on until $t$ exceeds $T_d$.
At that point, the policy is terminated, and a new episode can begin on a new image.

\begin{table}[h!]
\centering
\caption{Summary of the notation.}
\label{tab:notation}
\begin{tabular}{|l|l|}
  \hline
  $\mathcal{I}$ & image \\
  $C_k$         & presence of class $k \in \{1,\dots,K\}$ \\ 
  $t$           & time into episode \\ 
  $T_s$, $T_d$  & start and deadline times \\ 
  $s^j$         & belief state at step $j$ \\ 
  $\pi$         & policy function, $b \mapsto a \in \mathcal{A}$ \\
  $\mathcal{A}$ & set of actions $a$\\ 
  \comment{$\mathcal{F}$ & set of featurization actions \\}
  \comment{$\mathcal{L}$ & set of classification actions\\}
  $o_i$         & a real-valued observation upon executing $a_i \in \mathcal{A}$\\
  $\mathcal{O}$ & set of executed actions\\
  $\mathbf{o}$  & set of observations $\{o_i | a_i \in \mathcal{O}\}$\\
  $c(a_i)$        & cost of executing $a_i$, in units of $t$\\
  \hline
\end{tabular}\end{table}

\subsection{Selecting actions based on expected reward} \label{sec:value}
As our goal is to pick actions dynamically, we want a function $Q(s,a): S \times \mathcal{A} \mapsto \mathbb{R}$, where $S$ is the space of all possible states, to assign a value to a potential action, given the current state of the decision process.
We can then define the desired policy $\pi$ as simply taking the (untaken) action with the maximum value:
\begin{align}
\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} Q(s,a_i)
\end{align}

Although the action space $\mathcal{A}$ is quite manageable, consisting of the detectors and classifier we would like to run on the image, the space of possible states $B$ is infinite.
Therefore we cannot learn a tabular representation of $Q(s,a)$, and so use function approximation to represent it \cite{Sutton1998}.
We featurize the state-action pair and assume linear structure:
\begin{align}
Q^\pi(s,a_i) = \theta_\pi^\top  \phi(s,a_i)
\end{align}

The policy's performance at time $t$ is determined by the detections that are part of the set of observations $\mathbf{o}$ at the last state $s^j$ before $t$.
Detection results of unobserved classes are an empty set.
Therefore, the final AP vs. Time evaluation of the episode is a function of the history of execution $h^0=s^0,s^1,\dots,s^J$, where $J$ is the last step of the process with $t \le T_d$.
We call the final evaluation function $eval(h^0,T_s,T_d)$, and it is precisely the area under the AP vs. Time curve (normalized by total possible area) between $T_s$ and $T_d$, as determined by the detections in $\mathbf{o^j}$ for all steps $j$ in the episode $h$.

@@

Specifically, as shown in Figure~\ref{fig:rewards}, we define the \emph{reward} of an action as
\begin{align}\label{eq:advanced}
R(s^j,a_i) = \frac{\Delta \text{ap}_i (t_T^j-\frac{1}{2}\Delta t_i)}{(1-\text{ap}^j)t_T^j}
\end{align}
where $t_T^j$ and $\text{ap}^j$ are the time left until deadline and the AP at state $s^j$, and $\Delta t_i$ and $\Delta \text{ap}_i$ are the time taken and AP change produced by the action $a_i$.

@@

We can then represent the final evaluation in terms of individual rewards:
\begin{align}
eval(h^0) = \sum_{j=0}^J R(s_j,a^{j})
\end{align}

And we can represent the expected value of the final evaluation recursively in terms of the value function:
\begin{align} \label{eq:recursive_value}
Q(s^j,a_i) = \mathbb{E}_{s^{j+1}} [R(s^j,a_i) + \gamma Q(s^{j+1},\pi(s^{j+1}))]
\end{align}
where $\gamma \in [0,1]$ is a \emph{discount} value that can mitigate the effects of increasing state-transition uncertainty over long episodes.

\subsection{Learning the policy}
While we can't directly compute the expectation in \eqref{eq:recursive_value}, we can sample it by running actual episodes to gather $<s,a,r,s'>$ samples, where $r$ is the reward obtained by taking action $a$ in state $s$, and $s'$ is the following state.

\comment{
Specifically, we collect such samples to the end of the episode, and are thus able to compute the actual discounted sum of rewards for $Q(s^j,a_i)$:
\begin{align}
Q(s^j,a_i) = \theta_\pi^\top \phi(s,a_i) = \sum_{i=0}^{J-j} \lambda^i R(b_{j+i},a^{j+i})
\end{align}
}

Learning the policy is then a problem of repeatedly gathering samples with the current policy, minimizing the error between the discounted reward to the end of the episode as predicted by our current $Q(s^j,a_i)$ and the actual values gathered, and updating the policy with the resulting weights.
This is standard generalized policy iteration, or more specifically fitted Q-iteration \cite{Sutton1998,Ernst2005}.

We use $L_2$-regularized regression to minimize the error.
We run $15$ iterations of accumulating samples by running $350$ episodes, starting with a baseline policy which will be described in \autoref{sec:evaluation}, and cross-validating the regularization parameter at each iteration; samples are not thrown away between iterations.

One meta-parameter of the approach is the discount $\gamma$.
With $\gamma=0$, the value function is determined entirely by the immediate reward; learning in this case can only result in completely greedy policies.
With $\gamma=1$, the value function is determined by the expected rewards to the end of the episode, which is an approximation to the final evaluation metric.
However, the highest value for $\gamma$ is not necessarily best, due to the interplay between the expressive power of the value function and the actual state transition behavior of the world.
We experiment with several values of $\gamma$, and find a mid-level value of $0.4$ to work best.

\subsection{Feature representation}
Our policy is at its base determined by a linear function of the features of the state: $\pi(s) = \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \theta_\pi^\top \phi(s,a_i)$.
Since we want to be able to learn a dynamic policy, the observations $\mathbf{o}$ that are part of the state $s$ should play a role in determining the value of a potential action.
We include the following quantities as features $\phi(s,a)$:
\begin{description}
\item[$P(C_a)$] The prior probability of the class that corresponds to the detector of action $a$
\item[$P(C_0|\mathbf{o}) \ldots P(C_1|\mathbf{o})$] The probabilities of all classes, conditioned on the current set of observations.
\item[$H(C_0|\mathbf{o}) \ldots H(C_1|\mathbf{o})$] The entropies of all classes, conditioned on the current set of observations.
\end{description}

Additionally, we include the mean and maximum entropies of all classes, and time features that represent the times until start and deadline.
To formulate learning the policy as a single regression problem, we represent the features in block form, where $\phi(s,a_i)$ is a vector of size $F|\mathcal{A}|$, with $F=47$ for $20$ classes and the features as described, with all values set to $0$ except for the block corresponding to $a_i$.

Jumping ahead, we visualize the learned weights on these features in \autoref{fig:weights}, reshaped to a matrix of size $|\mathcal{A}| \times F$.

\begin{figure}[h!]
\centering
\includegraphics[width=0.87\linewidth]
    {../figures/weights.pdf}
\caption{Weights \todo{describe}}
\label{fig:weights}
\end{figure}

\subsection{Updating with observations}
\begin{figure}[h!]
\centering
\includegraphics[width=0.56\linewidth]{../figures/mrf_w_gist.pdf}
\caption{
The MRF inference model used in our system. The shaded nodes represent two actions that have already been taken; their observations have been recorded.
}
\label{fig:model}
\end{figure}
